---
title: "Activity Monitor Data Analysis"
author: "Stephen Kaluzny"
date: "07/11/2014"
output: html_document
---

## The Data
```{r}
require("dplyr")
activity <- tbl_df(read.csv(unz("activity.zip", file="activity.csv"),
    stringsAsFactors = FALSE))
activity$date <- as.Date(activity$date)
summary(activity)
```

## Steps Per Day
We sum the number of steps for each day. 
```{r}
require("ggplot2")
activity_by_date <- group_by(activity, date)
steps_per_day <- summarise(activity_by_date, total_steps=sum(steps, na.rm=TRUE))
# hist(steps_per_day$total_steps)
ggplot(steps_per_day, aes(x=total_steps)) + geom_histogram() +
    xlab("Total Number of Steps Per Day")
mean_steps_per_day <- mean(steps_per_day$total_steps)  # mean steps per day
median_steps_per_day <- median(steps_per_day$total_steps)  # median steps per day
```
The mean number of steps per day is `r format(mean_steps_per_day)`
and the median number of steps per day is
`r format(median_steps_per_day)`.

## Steps By Interval
```{r}
activity_by_interval <- group_by(activity, interval)
interval_average <- summarise(activity_by_interval, ave=mean(steps, na.rm=TRUE),
    med=median(steps, na.rm=TRUE))
ggplot(interval_average, aes(x=interval, y=ave)) +
    geom_line() +
    xlab("interval during a day") +
    ylab("Average Number of Steps")
```

## Missing Values

```{r}
nonmissing_steps_per_day <- summarise(activity_by_date,
    nonmissing=sum(!is.na(steps)))
table(nonmissing_steps_per_day$nonmissing)
```

The missing values occur only in the `steps` variable:
Missing date values: `r sum(is.na(activity$date))`,
missing interval values: `r sum(is.na(activity$interval))`,
missing steps values: `r sum(is.na(activity$steps))`,

We will replace any missing value in `steps` with the mean value
for `steps` across all days for the same time interval value.

Get the index (row number) where `steps` is missing.
Match the value of `interval` for theses missing rows
with the value of `interval` in the `interval_average` data frame.
Use the match index value to get the mean step value value to use
in place of the missing value.
```{r}
indxMissingSteps <- which(is.na(activity$steps))
activity_imputed <- activity
indxInterval <- match(activity[indxMissingSteps, "interval"],
   interval_average[, "interval"])
activity_imputed[indxMissingSteps, "steps"] <-
    interval_average[indxInterval, "ave"]
activity_imputed_by_date <- group_by(activity_imputed, date)
steps_imputed_per_day <- summarise(activity_imputed_by_date,
    total_steps=sum(steps, na.rm=TRUE))
ggplot(steps_imputed_per_day, aes(x=total_steps)) + geom_histogram()
imputed_mean_steps_per_day <- mean(steps_imputed_per_day$total_steps)
imputed_median_steps_per_day <- median(steps_imputed_per_day$total_steps)
```
The mean number of steps per day with imputed data is
`r format(imputed_mean_steps_per_day)`.
and the median number of steps per day with imputed data is
`r format(imputed_median_steps_per_day)`.

## Weekday Versus Weekend
```{r}
activity_imputed$daytype <- ifelse(weekdays(activity_imputed$date) %in%
    c("Saturday", "Sunday"), "weekend", "weekday")
table(activity_imputed$daytype)
```
```{r}
activity_imputed_by_interval_daytype <- group_by(activity_imputed, interval, daytype)
interval_average_by_daytype <- summarise(activity_imputed_by_interval_daytype, ave=mean(steps, na.rm=TRUE))
ggplot(interval_average_by_daytype, aes(x=interval, y=ave)) +
    geom_line() +
    facet_grid(daytype ~ .)
```
